EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-14 draft.tex
Some revision requests here and there

  \includegraphics[width=0.30\textwidth]{GAOnly_f005}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f005}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f005}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{
Scaling of running time with problem dimension to reach certain target values $\Delta f$. Lines:
average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
the most difficult target that was reached at least once (but not always);
Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
interquartile range with median of simulated runs; all values are divided by
dimension and plotted as $log_{10}$ values versus dimension. 
}
\label{fig:bbob}
\end{figure*}

@@

Having a multi-population based algorithm can decrease the execution time
because of the parallel execution of the evolution. But, having heterogeneous 
populations might enhance evolutionary search and needless
evaluations than homogeneous systems; heterogeneous settings, if
done right, increases the diversity of the whole population\cite{araujo2008multikulti}.

@@

Therefore, in this experiment we compare a multi-population with only GA and PSO populations,
versus an ensemble of GA and PSO algorithms. We tested on the first five functions of the
BBOB testbed. We used ten populations and eight workers for the experiment and the
same parameters as before.

@@

The results obtained by the experiments show (see Figure \ref{fig:bbob}) how
the runtime scales with dimension to reach certain target values $\Delta f$;
The color of the lines indicates the average runtime, for each target reached
at least one time. A red circle without a number on top indicates the algorithm
has reached the $10^{-8}$ target on all instances for that dimension. If there
is a number on the top, it indicates how many times the algorithm has reached
the target. Fixed values of targets $\Delta f = 10^{k}$ with $k$ colors are
given in the legend of $f_1$, $k$ values are $[-8,-5,-3,-2,-1,0,1]$. Results
from each configuration are presented in each column, GA, PSO, and GA\&PSO
populations.

We can see that the GA algorithm (on the left column) has a hard time reaching even the
$10^{-5}$ target in functions ($f_1-f_4$) for all dimensions, and has the worst
performance in 20 and 40 dimensions. Finally, it performs better on $f_5$
reaching the $10^{-8}$ on all dimensions but with a higher runtime on lower
dimensions.

On the other hand, the multi-population PSO (taking the column on the middle) 
has better performance; its results take a smaller range,
reaching all targets for functions $f_1$ and $f_2$ but with a slightly higher
runtime than the multi-strategy configuration. It is not able to reach the
$10^{-8}$ target on the higher dimensions of $f_3$ and $f_4$. Moreover, the PSO
multi-population has the best runtime for the lower dimensions of $f_5$.

Finally, the proposed PSO\&GA mixed algorithm has reached the $10^{-8}$ target
value on all functions ($f_1-f_5$) and scales well to higher dimensions, even on
$f_3$ and $f_4$ (charts on the lower rows), which are usually considered
difficult functions. This experiment also exemplifies the type of analyses that
we can perform with the EvoSwarm application by reusing the COCO framework.

@@

  \item[A virtual machine] with a container-optimized operating system, for example,
  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
  virtual machine and run docker-compose commands as we would do in a local 
  workstation. Salza et al. \cite{salza2019speed} tested this option by using
  CoreOs cluster. 
  % Remember that Salza et al. used CoreOS, which is optimized for containers - JJ

@@

% This is the second option. The first option was already tested, right? - JJ
% This is the third option.  - M

As a proof-of-concept, we chose the third option and deployed the EvoSwarm application in Amazon Elastic
Container Service (ECS),  since it needs fewer changes in configuration 
(from deploying it locally using docker-compose) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ

@@

For the second option, we used the same cluster and additionally
assigned each worker to a task with 4GBs of memory and 2048 computing units. Then scale the 
number of worker tasks to eight. We can scale the number of tasks for up to 100 instances in 
a Fargate configuration. 
                                  % Please use caps or not for Tasks in a consistent way - JJ
                                  % Done -Mario
                                  % Done above - M

\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
version: 1
task_definition:
  task_execution_role: ecsTaskExecutionRole
  ecs_network_mode: awsvpc
  task_size:
    mem_limit: 8GB
    cpu_limit: 4096
run_params:
  network_configuration:
    awsvpc_configuration:
      subnets:
        - "subnet-0ca2a542bbd74239e"
        - "subnet-0dfd367b5ca531b07"
      security_groups:
        - "sg-0681c18469874b391"
      assign_public_ip: ENABLED
\end{lstlisting}

@@

    controller:
      image: mariosky/controller
      volumes:
          - /code
      command: python controller.py
      environment:
          PYTHONUNBUFFERED: 1
          REDIS_HOST: localhost
      logging:
        driver: awslogs
        options:
          awslogs-region: us-west-2
          awslogs-group: docker-logs
          awslogs-stream-prefix: controller
  \end{lstlisting}

@@

% Please clarify that those images are published in Docker hub by yourself - JJ
The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.

>>>>

  \includegraphics[width=0.30\textwidth]{GAOnly_f005}&
  \includegraphics[width=0.30\textwidth]{PSOOnly_f005}&
  \includegraphics[width=0.30\textwidth]{GAPSO_f005}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{
Scaling of running time with problem dimension to reach certain target values $\Delta f$. Lines:
average runtime (aRT); Cross ($+$): median runtime of successful runs to reach
the most difficult target that was reached at least once (but not always);
Cross (\textcolor{red}{$\times$}): maximum number of f-evaluations in any trial. Notched boxes:
interquartile range with median of simulated runs; all values are divided by
dimension and plotted as $log_{10}$ values versus dimension.
% More explanation for those not familiar with BBOB - JJ
}
\label{fig:bbob}
\end{figure*}

@@

A multi-population based algorithm can decrease the execution time
because of parallel execution. But, having heterogeneous 
populations might enhance evolutionary search and avoid unneeded
evaluations more than homogeneous systems systems do; heterogeneous settings, if
done right, increases the diversity of the whole population\cite{araujo2008multikulti}.

@@

Therefore, in this experiment we compare a multi-population with  GA- and PSO- only populations,
versus an ensemble of GA and PSO algorithms. We have tested on the first five functions of the
BBOB testbed; we have used ten populations and eight workers for the experiment and the
same parameters as before.

@@

The results obtained by the experiments show (see Figure \ref{fig:bbob}) how
the runtime scales with dimension to reach certain target values $\Delta f$;
The color of the lines indicates the average runtime, for each target reached
at least once. A red circle without a number on top indicates the algorithm
has reached the $10^{-8}$ target on all instances for that dimension. If there
is that number, it indicates how many times the algorithm has reached
the target. Fixed values of targets $\Delta f = 10^{k}$ with $k$ colors are
given in the legend of $f_1$, $k$ values are $[-8,-5,-3,-2,-1,0,1]$. Results
from each configuration are presented in each column, GA, PSO, and GA\&PSO
populations. % clarify which colors for every level - JJ

We can see that the GA algorithm (on the leftmost column) has a hard time reaching even the
$10^{-5}$ target in functions ($f_1-f_4$) for all dimensions, and has the worst
performance in 20 and 40 dimensions. Finally, it performs better on $f_5$
reaching the $10^{-8}$ on all dimensions but with a higher runtime on lower
dimensions.

On the other hand, the multi-population PSO (taking the column on the middle) 
has better performance; its results are spread over a smaller range,
reaching all targets for functions $f_1$ and $f_2$ but with a slightly higher
runtime than the multi-strategy configuration. It is not able to reach the
$10^{-8}$ target on the higher dimensions of $f_3$ and $f_4$. Moreover, the PSO
multi-population has the best runtime for the lower dimensions of $f_5$.

Finally, the proposed PSO\&GA mixed algorithm has reached the $10^{-8}$ target
value on all functions ($f_1-f_5$) and scales well to higher dimensions, even on
$f_3$ and $f_4$ (charts on the lower rows), which are usually considered
difficult functions. This experiment also exemplifies the type of analyses that
we can perform with the results of the EvoSwarm application by reusing the COCO framework. % If this is what you want to prove, it's not a very strong point. Maybe it would be stronger if you proved that these mixed algorithms can be tested in the same conditions as single-algorithm, and also that these mixed algorithms work better than single-algorithm versions - JJ

@@

  \item[A virtual machine] with a container-optimized operating system, for example,
  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
  virtual machine and run docker-compose commands as we would do in a local 
  workstation. Salza et al. \cite{salza2019speed} tested this option by using
  CoreOs cluster.

@@

As a proof-of-concept, we chose the third option, that is, a container orchestration service, and deployed the EvoSwarm application in Amazon Elastic
Container Service (ECS),  since it needs fewer changes in configuration 
(from deploying it locally using docker-compose) and it is has a lower cost than the Kubernetes alternative. % Remember to mention cost at the end - JJ

@@

For the second option, we used the same cluster and additionally
assigned each worker to a task with 4GBs of memory and 2048 computing units. Then scale the 
number of worker tasks to eight. We can scale the number of tasks for up to 100 instances in 
a Fargate configuration.

\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
version: 1
task_definition:
  task_execution_role: ecsTaskExecutionRole
  ecs_network_mode: awsvpc
  task_size:
    mem_limit: 8GB
    cpu_limit: 4096
run_params:
  network_configuration:
    awsvpc_configuration:
      subnets:
        - "subnet-0ca2a542bbd74239e"
        - "subnet-0dfd367b5ca531b07"
      security_groups:
        - "sg-0681c18469874b391"
      assign_public_ip: ENABLED
\end{lstlisting}
% Do you need the details about the subnets? As long as they are here, you'll have to explain them. Why two subnets? It's probably better to leave that network configuration out anyway - JJ

% Please clarify that those images are published in Docker hub by yourself - JJ

@@

    controller:
      image: mariosky/controller
      volumes:
          - /code
      command: python controller.py
      environment:
          PYTHONUNBUFFERED: 1
          REDIS_HOST: localhost
      logging:
        driver: awslogs
        options:
          awslogs-region: us-west-2
          awslogs-group: docker-logs
          awslogs-stream-prefix: controller
  \end{lstlisting}
% You will probably need to explain this a bit - JJ

@@

The results are shown in Figure~\ref{fig:aws-results} and Table~\ref{tab:aws}.
